Exploring dimensions of biodiversity in Japanese ferns

### 2021-05-18 ms/manuscript.Rmd
Update MS

**Methods** We used a list of >300,000 fern specimens in Japan to compile a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` native, non-hybrid taxa. We combined this with a phylogenetic tree and functional trait data to analyze taxonomic, phylogenetic, and functional diversity, and assessed significance using a null distribution of 999 random communities. We conducted categorical analysis of neo- and paleo-endemism to identify areas of significant phylogenetic endemism. We used hierarchical clustering of taxonomic and phylogenetic distances to analyze bioregions. We compared the overlap of significantly diverse grid-cells with existing conservation zones to assess conservation status.

@@

Presumably, the main reason phylogeny has not been taken into account more prominently in biodiversity studies is because data (DNA sequences) and analytical tools have only become available relatively recently. These tools and datasets now make it possible to analyze other dimensions of biodiversity, such as phylogenetic diversity and phylogenetic endemism. Better ways to characterize biological regions ("bioregions") are also available, rather than relying on ad-hoc characterizations. Furthermore, incorporating these two frameworks---the categorization of areas into phyloregions with analysis of over/under dispersion of biodiversity---can provide particularly powerful insights into the processes structuring biodiversity and suggest conservation priorities.

@@

Japan is situated along a latitudinal gradient spanning highly season temperate areas in the north to subtropical, mostly aseasonal islands in the south. Furthermore, it is a mountainous country with great variation in elevation on the larger islands. Plant distributions often reflect physiological adaptations to climate (e.g., percipiation and temperature). Therefore, we expect the spatial distribution of biodviversity to be determined by climatic variation in addition to reproductive mode.

@@

Our data include all ferns known to occur within the political boundaries of Japan (20°N to 45°N, 122°E to 153°E). Japan consists of four main islands and thousands of smaller ones (`r figure("japan-map")`). The main islands have been in contact with the continent at various points in the past during periods of lower sea level, but the Ryukyu and Ogasawara archipelagos are oceanic islands that have never been connected to the mainland. The climate is subtropical in the south and highly seasonal in other parts of the country. Map data were downloaded from the Geospatial Information Authority of Japan under the Creative Commons Attribution License v4.0 (https://www.gsi.go.jp/kankyochiri/gm_japan_e.html).

@@

A necessary step in any analysis of biodiversity is to set the grain size used to define communities. Here, one must consider that smaller grain size is needed to detect environmental effects, while larger grain size is needed to ensure adequate species sampling. We tested using 0.1°, 0.2°, 0.3°, and 0.4° grid-cells (size refers to length of side of a grid-cell) for defining communities (co-occurring species). At each grain size, species occurrences were converted into a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid-cell). We calculated sampling redundance (1 - richness/number of specimens) [@Garcillan2003] to quantify adequacy of sampling. Preliminary analysis indicated that 0.2° grid-cells are optimal for our dataset (`r s_figure("grain-size")`). We filtered out any cells with redundancy < 0.1 and used this for subsequent analyses.

@@

We used traits originally compiled for identification of ferns and lycophytes of Japan as follows [@Ebihara2016b; @Ebihara2017]. Continuous traits were measured on 10 randomly selected specimens per species. Fertile (\ie, spore-bearing) and sterile fronds were measured separately, then averaged. Qualitative traits were scored by observing voucher specimens. All qualitative traits were scored in binary format; \eg, a single trait with three states was formatted as three binary traits. From the original trait list, we selected only traits with putatative ecological function (`r table("traits-used")`), and removed any that had Pearson correlation coefficient >0.6 or fewer than three observations of a given trait state.

@@

We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `getData` function in the R package `raster`, and extracted four variables related to our research questions as follows: mean annual temperature (bio1), temperature seasonality (bio4), annual precipitation (bio12), and precipitation seasonality (bio15). We checked for correlation between climatic variables while taking into account spatial autocorrelation using a modified *t*-test with the `modified.ttest` in the `SpatialPack` R package; as a result, we excluded temperature seasonality, which was correlated with mean annual temperature. We identified the intersection of the climate data at the 2.5 minute scale with the 0.2° grid-cells, then calculated the mean of each climatic variable for each 0.2° grid-cell. The R package 'sf' was used for GIS analyses.

@@

We tested for a relationship between SES of each biodiversity metric and reproductive mode (% apomictic taxa) initially with linear models constructed using the R function 'lm'. Next, we checked for evidence of spatial autocorrelation in the linear model residuals with Moran's I using 'moran.mc' function in the spdep package. Spatial autocorrelatoin was frequently detected, so to acccount for this we constructed spatial linear mixed models including a Matérn correlation matrix based on the geographic centroids of each grid cell with the R package spaMM [@Rousset2014]. We then again checked for any remaining spatial autocorrelation in the mixed models with Moran's I.

@@

>>>>

**Methods** We used a list of >300,000 fern specimens in Japan to compile a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` native, non-hybrid taxa. We combined this with a phylogenetic tree and functional trait data to analyze taxonomic, phylogenetic, and functional diversity, and modeled the distribution of biodiversity in response to environmental factors and reproductive mode. We conducted categorical analysis of neo- and paleo-endemism to identify areas of significant phylogenetic endemism. We used hierarchical clustering of taxonomic and phylogenetic distances to analyze bioregions. We compared the overlap of significantly diverse grid-cells with existing conservation zones to assess conservation status.

@@

Presumably, the main reason phylogeny has not been taken into account more prominently in biodiversity studies is because data (DNA sequences) and analytic tools have only become available relatively recently. These tools and datasets now make it possible to analyze other dimensions of biodiversity, such as phylogenetic diversity and phylogenetic endemism. Better ways to characterize biological regions ("bioregions") are also available, rather than relying on ad-hoc characterizations. Furthermore, incorporating these two frameworks---the categorization of areas into phyloregions with analysis of over/under dispersion of biodiversity---can provide particularly powerful insights into the processes structuring biodiversity and suggest conservation priorities.

@@

Japan is situated along a latitudinal gradient spanning seasonal, temperate areas in the north to subtropical, mostly aseasonal islands in the south. Furthermore, it is a mountainous country with great variation in elevation on the larger islands. Plant distributions often reflect physiological adaptations to climate (e.g., precipitation and temperature). Therefore, we expect the spatial distribution of biodiversity to be determined by climatic variation in addition to reproductive mode.

@@

Our data include all ferns known to occur within the political boundaries of Japan (20°N to 45°N, 122°E to 153°E). Japan consists of four main islands and thousands of smaller ones (`r figure("japan-map")`). The main islands have been in contact with the continent at various points in the past during periods of lower sea level, but the Ryukyu and Ogasawara archipelagos are oceanic islands that have never been connected to the mainland. The climate is subtropical in the south and temperate in other parts of the country. Map data were downloaded from the Geospatial Information Authority of Japan under the Creative Commons Attribution License v4.0 (https://www.gsi.go.jp/kankyochiri/gm_japan_e.html).

@@

A necessary step in any analysis of biodiversity is to set the grain size used to define communities. Here, one must consider that smaller grain size is needed to detect environmental effects, while larger grain size is needed to ensure adequate species sampling. We tested using 0.1°, 0.2°, 0.3°, and 0.4° grid-cells (size refers to length of side of a grid-cell) for defining communities (co-occurring species). At each grain size, species occurrences were converted to a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid-cell). We calculated sampling redundancy (1 - richness/number of specimens) [@Garcillan2003] to quantify adequacy of sampling. Preliminary analysis indicated that 0.2° grid-cells are optimal for our dataset (`r s_figure("grain-size")`). We filtered out any cells with redundancy < 0.1 and used this for subsequent analyses.

@@

We used traits originally compiled for identification of ferns and lycophytes of Japan as follows [@Ebihara2016b; @Ebihara2017]. Continuous traits were measured on 10 randomly selected specimens per species. Fertile (\ie, spore-bearing) and sterile fronds were measured separately, then averaged. Qualitative traits were scored by observing voucher specimens. All qualitative traits were scored in binary format; \eg, a single trait with three states was formatted as three binary traits. From the original trait list, we selected only traits with putative ecological function (`r table("traits-used")`), and removed any that had Pearson correlation coefficient >0.6 or fewer than three observations of a given trait state.

@@

We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `getData` function in the R package `raster`, and extracted four variables related to our research questions as follows: mean annual temperature (bio1; hereafter "temperature"), temperature seasonality (bio4), annual precipitation (bio12; hereafter "precipitation"), and precipitation seasonality (bio15). We identified the intersection of the climate data at the 2.5 minute scale with the 0.2° grid-cells, then calculated the mean of each climatic variable for each 0.2° grid-cell. The R package 'sf' was used for GIS analyses.

@@

@@

### Spatial models

We first tested for spatial autocorrelation in biodiversity metrics and independent variables (environmental variables and % apomictic taxa) using Moran's I with the `moran.mc` function in the `spdep` package. In all cases, significant spatial autocorrelation was detected (Moran's I > 0, *P* value of permutation test < 0.05). Therefore, we used spatially explicit methods for modeling.

We checked for correlation between independent variables while taking into account spatial autocorrelation using a modified *t*-test with the `modified.ttest` in the `SpatialPack` R package. This indicated that temperature seasonality and % apomictic taxa were correlated with mean annual temperature. Therefore, we excluded temperature seasonality from modeling. We suspected that despite the correlation, % apomictic taxa might explain the distribution of genetic diversity better than temperature, so we conducted analyses separately for this variable as described below.

We constructed a set of linear models for each response variable (richness, SES of PD, SES of FD, SES of RPD, SES of RFD, PE p-score) as dependent on all possible combinations of environmental variables (temperature, precipitation, and precipitation seasonality). Initial inspection of the relationship between richness and temperature showed a hump-shaped pattern, so we also included a quadratic term for temperature for the richness model only. We fit models including a Matérn correlation matrix based on the geographic centroids of each grid cell with the R package spaMM [@Rousset2014], then identified the best-fitting model for each response variable as the one with the greatest log-likelihood. We verified the significance of each term of the best model by conducting likelihood ratio tests (LRTs) between the full model and a model with the term of interest removed, in sequence. We also verified the significance of the full model relative to a null model (only including the Matérn correlation matrix) using LRTs.

To compare the model fit between using temperature (environmental model) or % apomictic taxa (reproductive model) in response variables with a genetic component (SES of PD, SES of RPD, PE p-score), we constructed additional models where % apomictic taxa was used in place of temperature in the best-fitting model for each response variable. We then conducted LRTs between the environmental model vs. the reproductive model, as well as between the full reproductive model and models with each term of interest removed, and between the full reproductive model vs. the null model.
